Verifying Spatial Properties of Array Computations

### 2017-04-15 oopsla17/study.tex
Notes on hypothesis 2

\noindent
We revist each hypothesis and show the related data and conclusions.
%
\begin{enumerate}
\item Loops over arrays mainly read from arrays with a static pattern
based on constant offsets from (base or dervied) induction variables;
\end{enumerate}
%
We identified $108,773$ instances of assignments within loops in
which an array subscript flows to the right-hand side. We refer
to each one of these as an \emph{array computation}.
We performed the classification described above
and grouped the data by the right-hand side classification only:
%%
\begin{center}
\begin{tabular}{lS[table-format=4]@{\extracolsep{5pt}}S[table-format=2.2]c}
\textbf{RHS classification} & \textbf{Number} & \% &
\textit{grouping 1} \\ \hline
Affine                          & 34        & 0.03\%
& \multirow{3}{*}{81.43\%} \\ \cline{1-3}
Neighbourhood                   & 65334     & 60.06 \%  \\ \cline{1-3}
Neighbourhood + constants       & 23209     & 21.34\%  \\ \hline
Other                           & 20196     & 18.57 \%  \\ \hline \hline
Total                           & 108773    &  \\
\end{tabular}
\end{center}
%
\noindent
Since, affine and neighbour patterns are static, we
group them (\textit{grouping 1}) to see that $81.43\%$ of the array
computations (88,577) have a static pattern that are either constants or
constant offsets of base/derived induction variables, confirming
\emph{hypothesis 1}. Note that affine (derived induction variables)
are rare. Overwhelmingly, the main category is neighbourhood
offsets (possibly mixed with constants in some
dimensions). The ``affine + constants'' class is
not represented at all.
%
\begin{enumerate}[resume]
\item Most loop-array computations of the previous form read
from arrays with a \emph{contiguous} pattern, \eg{}:
%
\begin{minted}{fortran}
do i = 1, n
  b(n) = (a(i) + a(i+1)) / 2.0
\end{minted}
%
\item Most loop-array computations of the previous form read
from arrays with a pattern that includes the immediate
neighbours (offsets of 1 from the induction variables);
\end{enumerate}
%
We grouped the data based on subclassifications of \emph{grouping 1}
in terms of contiguity and positioning.
%Focussing on grouping 1.
\begin{center}
\begin{tabular}{lS[table-format=4]S[table-format=2.2]S[table-format=2.2]lS[table-format=4]S[table-format=2.2]S[table-format=2.2]}
\textbf{RHS classification} & \textbf{Number} & \% {(of whole)}
                                              & \% {(of group)}
&\textbf{Position} & \textbf{Number} & \% {(whole)} & \% {(group)} \\ \hline
Affine, contiguous
& 31 & 0.03\% & 0.03\% \\ \hline
%\multirow{3}{*}{Affine, contiguous}
% & %\multirow{3}{*}{31}
% & %\multirow{3}{*}{0.03\%}
% & \textsf{away} & 4 & 0.00\% \\
%& 31 & 0.03\% & \textsf{origin} & 15 & 0.01\% \\
%&&& \textsf{straddle} & 12 & 0.01\% \\ \hline
Affine, non-contiguous
 & 3
 & 0.00\% & 0.00\% \\ \hline
\multirow{3}{*}{Neighbour, contiguous}
 & %\multirow{3}{*}{83689}
 & %\multirow{3}{*}{83.08\%}
& & \textsf{away} & 152 & 0.14\% & 0.17 \% \\
& 83689 & 76.94\% & 94.48\% & \textsf{origin} & 81309 & 74.75\% & 91.79\% \\
&&&& \textsf{straddle} & 2228 & 2.05\% & 2.52\% \\ \hline
Neighbour, non-contiguous
 & 4854
 & 4.46\% & 5.48\%  \\ \hline \hline
\textbf{Total}  & 88577    &  \\
\end{tabular}
\end{center}
%%
Thus, $76.94\%$ of array computations have only contiguous
neighbour indices (which may include constants in some indices)
in a contiguous pattern on the right hand side, with $76.80\%$
either at the origin ($74.75\%$) or offset by a distance of 1 from the origin
($2.05\%$). This confirms both hypotheses 2 and 3. Only 152 instances
were found of index patterns not immediate next to or over the origin.

>>>>

\noindent
We revist each hypothesis and show the related data and conclusions.
%
\begin{enumerate}
\item Loops over arrays mainly read from arrays with a static pattern
based on constant offsets from (base or dervied) induction variables;
\end{enumerate}
%
We identified $108,773$ instances of assignments within loops in
which an array subscript flows to the right-hand side. We refer
to each one of these as an \emph{array computation}.
We performed the classification described above
and grouped the data by the right-hand side classification only:
%%
\begin{center}
\begin{tabular}{lS[table-format=4]@{\extracolsep{5pt}}S[table-format=2.2]c}
\textbf{RHS classification} & \textbf{Number} & \% &
\textit{grouping 1} \\ \hline
Affine                          & 34        & 0.03\%
& \multirow{3}{*}{81.43\%} \\ \cline{1-3}
Neighbourhood                   & 65334     & 60.06 \%  \\ \cline{1-3}
Neighbourhood + constants       & 23209     & 21.34\%  \\ \hline
Other                           & 20196     & 18.57 \%  \\ \hline \hline
Total                           & 108773    &  \\
\end{tabular}
\end{center}
%
\noindent
Since, affine and neighbour patterns are static, we
group them (\textit{grouping 1}) to see that $81.43\%$ of the array
computations (88,577) have a static pattern that are either constants or
constant offsets of base/derived induction variables, confirming
\emph{hypothesis 1}. Note that affine (derived induction variables)
are rare. Overwhelmingly, the main category is neighbourhood
offsets (possibly mixed with constants in some
dimensions). The ``affine + constants'' class is
not represented at all.
\mnote{(1) State that constant here includes variables that are constant with
respect to the current loop scope.}
\mnote{(2) Mention what sort of things are
included in other.}
\mnote{(3) Add a remark about lack of affine indices to be
perhaps programmers relying on compiler more regarding optimisations than hand
optimising. We can later use this to justify our inability to handle affine
indices.}
%
\begin{enumerate}[resume]
\item Most loop-array computations of the previous form read
from arrays with a \emph{contiguous} pattern, \eg{}:
%
\begin{minted}{fortran}
do i = 1, n
  b(n) = (a(i) + a(i+1)) / 2.0
\end{minted}
%
\item Most loop-array computations of the previous form read
from arrays with a pattern that includes the immediate
neighbours (offsets of 1 from the induction variables);
\end{enumerate}
%
We grouped the data based on subclassifications of \emph{grouping 1}
in terms of contiguity and positioning.
%Focussing on grouping 1.
\begin{center}
\begin{tabular}{lS[table-format=4]S[table-format=2.2]S[table-format=2.2]lS[table-format=4]S[table-format=2.2]S[table-format=2.2]}
\textbf{RHS classification} & \textbf{Number} & \% {(of whole)}
                                              & \% {(of group)}
&\textbf{Position} & \textbf{Number} & \% {(whole)} & \% {(group)} \\ \hline
Affine, contiguous
& 31 & 0.03\% & 0.03\% \\ \hline
%\multirow{3}{*}{Affine, contiguous}
% & %\multirow{3}{*}{31}
% & %\multirow{3}{*}{0.03\%}
% & \textsf{away} & 4 & 0.00\% \\
%& 31 & 0.03\% & \textsf{origin} & 15 & 0.01\% \\
%&&& \textsf{straddle} & 12 & 0.01\% \\ \hline
Affine, non-contiguous
 & 3
 & 0.00\% & 0.00\% \\ \hline
\multirow{3}{*}{Neighbour, contiguous}
 & %\multirow{3}{*}{83689}
 & %\multirow{3}{*}{83.08\%}
& & \textsf{away} & 152 & 0.14\% & 0.17 \% \\
& 83689 & 76.94\% & 94.48\% & \textsf{origin} & 81309 & 74.75\% & 91.79\% \\
&&&& \textsf{straddle} & 2228 & 2.05\% & 2.52\% \\ \hline
Neighbour, non-contiguous
 & 4854
 & 4.46\% & 5.48\%  \\ \hline \hline
\textbf{Total}  & 88577    &  \\
\end{tabular}
\end{center}
%%
Thus, $76.94\%$ of array computations have only contiguous
neighbour indices (which may include constants in some indices)
in a contiguous pattern on the right hand side, with $76.80\%$
either at the origin ($74.75\%$) or offset by a distance of 1 from the origin
($2.05\%$). This confirms both hypotheses 2 and 3. Only 152 instances
were found of index patterns not immediate next to or over the origin.
